Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2023-10-23 paper/main.tex
backing up progress (working on addressing R2 comments)

We adapted an approach we developed in prior work~\citep{HeusEtal21} to embed
each moment of the two lectures and each question in our pool in a common
representational space. Briefly, our approach uses a topic model (Latent
Dirichlet Allocation;~\citealp{BleiEtal03}), trained on a set of documents, to
discover a set of (up to) $k$ ``topics'' or ``themes.'' Formally, each topic is
defined as a distribution of weights over each word in the model's vocabulary
(i.e., the union of all unique words, across all documents, excluding ``stop
words.''). Conceptually, each topic is intended to give larger weights to words
that are semantically related or tend to co-occur in the same documents. After
fitting a topic model, each document in the training set, or any \textit{new}
document that contains at least some of the words in the model's vocabulary,
may be represented as a $k$-dimensional vector describing how much the document
(most probably) reflects each topic. To select an appropriate $k$ for our
model, we identified the minimum number of topics that yielded at least one
``unused'' topic (i.e., in which all words in the vocabulary were assigned
uniform weights) after training. This indicated that the number of topics was
sufficient to capture the set of latent themes present in the two lectures
(from which we constructed our document corpus, as described below). We found
this value to be $k = 15$ topics. The distribution of weights over words in the
vocabulary for each discovered topic is shown in Supplementary
Figure~\topicWordWeights, and each topic's top-weighted words may be found in
Supplementary Table~\topics.

@@

We fetched these automated transcripts using the
\texttt{youtube-transcript-api} Python package~\citep{Depo18}. The transcripts
consisted of one timestamped line of text for every few seconds (mean: 2.34~s;
standard deviation: 0.83~s) of spoken content in the video (i.e., corresponding
to each individual caption that would appear on-screen if viewing the lecture
via YouTube, and when those lines would appear). After scraping the
transcripts, we carried out several standard preprocessing steps. First, we
removed all punctuation, replaced numbers with their word equivalents (e.g.,
``1'' with ``one''), and converted all text to lowercase. Next, we removed all
stop words, as defined by the Natural Language Toolkit~\citep[NLTK;
]{BirdEtal09} English stop word list, augmented with the following additional
words that we hand-identified as frequently occuring and non-diagnostic:
``actual,'' ``actually,'' ``also,'' ``bit,'' ``could,'' ``e,'' ``even,''
``first,'' ``follow,'' ``following,'' ``four,'' ``let,'' ``like,'' ``mc,''
``really,'', ``saw,'' ``see,'' ``seen,'' ``thing,'' and ``two.'' We defined a
sliding window length of (up to) $w = 30$ transcript lines, and assigned each
window a timestamp corresponding to the midpoint between the timestamps for its
first and last lines. This $w$ parameter was chosen to match the same number of
words per sliding window (rounded to the nearest whole word, and after
preprocessing) as the sliding windows we defined in our prior
work~\citep{HeusEtal21}. The sliding windows contained an average of 73.8 words
and lasted for an average of 62.22 seconds.

>>>>

We adapted an approach we developed in prior work~\citep{HeusEtal21} to embed
each moment of the two lectures and each question in our pool in a common
representational space. Briefly, our approach uses a topic model (Latent
Dirichlet Allocation;~\citealp{BleiEtal03}), trained on a set of documents, to
discover a set of (up to) $k$ ``topics'' or ``themes.'' Formally, each topic is
defined as a distribution of weights over each word in the model's vocabulary
(i.e., the union of all unique words, across all documents, excluding ``stop
words.''). Conceptually, each topic is intended to give larger weights to words
that are semantically related, as implied by their co-occurring in the same documents. After
fitting a topic model, each document in the training set, or any \textit{new}
document that contains at least some of the words in the model's vocabulary,
may be represented as a $k$-dimensional vector describing how much the document
(most probably) reflects each topic. To select an appropriate $k$ for our
model, we identified the minimum number of topics that yielded at least one
``unused'' topic (i.e., in which all words in the vocabulary were assigned
uniform weights) after training. This indicated that the number of topics was
sufficient to capture the set of latent themes present in the two lectures
(from which we constructed our document corpus, as described below). We found
this value to be $k = 15$ topics. The distribution of weights over words in the
vocabulary for each discovered topic is shown in Supplementary
Figure~\topicWordWeights, and each topic's top-weighted words may be found in
Supplementary Table~\topics.

@@

We fetched these automated transcripts using the
\texttt{youtube-transcript-api} Python package~\citep{Depo18}. The transcripts
consisted of one timestamped line of text for every few seconds (mean: 2.34~s;
standard deviation: 0.83~s) of spoken content in the video (i.e., corresponding
to each individual caption that would appear on-screen if viewing the lecture
via YouTube, and when those lines would appear). We defined a sliding window
length of (up to) $w = 30$ transcript lines, and assigned each window a
timestamp corresponding to the midpoint between the timestamps for its first
and last lines. This $w$ parameter was chosen to match the same number of words
per sliding window (rounded to the nearest whole word, and after preprocessing)
as the sliding windows we defined in our prior work~\citep{HeusEtal21}. The
sliding windows contained an average of 73.8 words and lasted for an average of
62.22 seconds. After splitting the transcripts into sliding windows, we carried
out several standard preprocessing steps. First, we removed all punctuation,
replaced numbers with their word equivalents (e.g., ``1'' with ``one''), and
converted all text to lowercase. Next, we removed all stop words, as defined by
the Natural Language Toolkit~\citep[NLTK; ]{BirdEtal09} English stop word list,
augmented with additional words following~\citep{BoydMimn14}: ``actual,''
``actually,'' ``also,'' ``bit,'' ``could,'' ``e,'' ``even,'' ``first,''
``follow,'' ``following,'' ``four,'' ``let,'' ``like,'' ``mc,'' ``really,'',
``saw,'' ``see,'' ``seen,'' ``thing,'' and ``two.''
